---
title: "PA1_template.Rmd"
author: "Ilmur Jonsdottir"
date: "5/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## First I set up the environment

```{r}
library("readr")
library("tidyr")
library("dplyr")
library("ggplot2")
library("VIM")
library("mice")
set.seed(1010)
```


##1. Code for reading in the dataset and/or processing the data

```{r}
activity <- read_csv("activity.csv")
```

##2. Histogram of the total number of steps taken each day

```{r}
activity %>% group_by(date) %>% summarise(stepsPerDay = sum(steps)) %>% 
        ggplot(aes(x = stepsPerDay)) + geom_histogram(bins = 15) + 
        ggtitle("Histogram: Total Number of Steps Per Day") + xlab("Steps Per Day") + 
        ylab("Frequency")
```

##3. Mean and median number of steps taken each day

```{r}
activity %>% group_by(date) %>% 
        summarise(meanStepsPerDay = mean(steps, na.rm = TRUE)) %>% 
        summarise(meanSteps = mean(meanStepsPerDay, na.rm = TRUE))

activity %>% group_by(date) %>% 
        summarise(medianStepsPerDay = median(steps, na.rm = TRUE)) %>%
        summarise(medianSteps = median(medianStepsPerDay, na.rm = TRUE))
```

##4.Time series plot of the average number of steps taken

```{r}
activity %>% group_by(date) %>% summarise(meanSteps = mean(steps, na.rm = TRUE)) %>%
        ggplot(aes(x = date, y = meanSteps)) + geom_line() + 
        ggtitle("Mean Steps by Date") + xlab("Date") + ylab("Mean Steps")
```

##5. The 5-minute interval that, on average, contains the maximum number of steps

```{r}
activity %>% group_by(interval) %>% 
        summarize(meanByInterval = mean(steps, na.rm = TRUE)) %>%
        filter(meanByInterval == max(meanByInterval))

activity %>% group_by(interval) %>% 
        summarize(meanByInterval = mean(steps, na.rm = TRUE)) %>%
        filter(meanByInterval == min(meanByInterval))
```

##6. Code to describe and show a strategy for imputing missing data

```{r}
md.pattern(activity)

(missing <- sum(is.na(activity)))

missingPercent <- sum(is.na(activity))/(dim(activity)[1]*dim(activity)[2]) * 100 
pMiss <- function(x) { sum(is.na(x)) / length(x) * 100}
(missingPercentCol <-apply(activity, 2, pMiss))

aggr(activity, numbers = TRUE)

activityNoMissing <- activity[complete.cases(activity),]
```

##7. Histogram of the total number of steps taken each day after missing values are removed

```{r}
activityNoMissing %>% group_by(date) %>% summarise(stepsPerDay = sum(steps)) %>% 
        ggplot(aes(x = stepsPerDay)) + geom_histogram(bins = 15) + 
        ggtitle("Histogram: Total Number of Steps Per Day") + xlab("Steps Per Day") + 
        ylab("Frequency")
```

##7a. Time series plot of average number of steps taken after missing values are removed

```{r}
activityNoMissing %>% group_by(date) %>% summarise(meanSteps = mean(steps, na.rm = TRUE)) %>%
        ggplot(aes(x = date, y = meanSteps)) + geom_line() + 
        ggtitle("Mean Steps by Date") + xlab("Date") + ylab("Mean Steps")
```

##8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
t <- activityNoMissing %>% mutate(dayOfWeek = weekdays(date)) %>%
        mutate(Weekend = ifelse(dayOfWeek == "Saturday" | dayOfWeek == "Sunday", "Weekend", "Weekday"))

t %>% 
        group_by(Weekend, interval) %>% mutate(meanStepsInterval = mean(steps)) %>%
        ggplot(aes(x = interval, y = meanStepsInterval)) + geom_line() +
        facet_wrap(~Weekend) +ggtitle("Mean Steps by Interval: Weekday vs. Weekend") + 
        xlab("Interval") + ylab("Mean Steps")

t %>%
        group_by(dayOfWeek, interval) %>% mutate(meanStepsInterval = mean(steps)) %>%
        ggplot(aes(x = interval, y = meanStepsInterval)) + geom_line() +
        facet_wrap(~dayOfWeek) +ggtitle("Mean Steps by Interval: By Day") +
        xlab("Interval") + ylab("Mean Steps")
```

